FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2017-10-20 NewVersion/sections/Auxiliary.tex
spellcheck, grammar

% \section{Method Lookup Algorithm and Auxiliary Definitions}\label{sec:auxdefs}
\section{Key Algorithms and Type-Soundness}\label{sec:auxdefs}
In this section, we present the fundamental algorithms and auxiliary
definitions used in our formalization, and show that the resulting
calculus is type sound. The functions presented in this section are
the key components that implement our algorithm for method lookup.

@@

\saveSpaceFig
\begin{flalign*}
	& \rhd \textit{Definition of } \mostSpecific(m, I, J): & \\
	& \bullet \mostSpecific(m, I, J) = \prune(origins) & \\
	& \indent\indent \textrm{with: } origins = \{K \mid \subt{I}{K}, \textrm{ and } \subt{K}{J} \; \lor \; \subt{J}{K}, &\\
	& \hspace{1.62in} \textrm{ and } K[m\ \kwoverride\ K] \textrm{ is defined} \} &
\end{flalign*}
By the definition, an interface belongs to $\mostSpecific(m, I, J)$ if and only if:
\begin{itemize}
	\item It originally defines $m$;
	\item It is a super type of $I$ (including $I$);
	\item It is either a super type or a subtype of $J$ (including $J$);
	\item Any subtype of it does not belong to the same result set because of $\prune$.
\end{itemize}
\saveSpaceFig

@@

The $\prune$ function takes a set of
types, and filters out those that have subtypes in the same set. In the returned set,
none of them has subtyping relation to one another, since all super types have been removed.
%%%%============================ prune(set) end ================%%%%%%%%

>>>>

% \section{Method Lookup Algorithm and Auxiliary Definitions}\label{sec:auxdefs}
\section{Key Algorithms and Type-Soundness}\label{sec:auxdefs}
In this section, we present the fundamental algorithms and auxiliary
definitions used in our formalization and show that the resulting
calculus is type sound. The functions presented in this section are
the key components that implement our algorithm for method lookup.

@@

\saveSpaceFig
\begin{flalign*}
	& \rhd \textit{Definition of } \mostSpecific(m, I, J): & \\
	& \bullet \mostSpecific(m, I, J) = \prune(origins) & \\
	& \indent\indent \textrm{with: } origins = \{K \mid \subt{I}{K}, \textrm{ and } \subt{K}{J} \; \lor \; \subt{J}{K}, &\\
	& \hspace{1.62in} \textrm{ and } K[m\ \kwoverride\ K] \textrm{ is defined} \} &
\end{flalign*}
By the definition, an interface belongs to $\mostSpecific(m, I, J)$ if and only if:
\begin{itemize}
	\item It originally defines $m$;
	\item It is a supertype of $I$ (including $I$);
	\item It is either a supertype or a subtype of $J$ (including $J$);
	\item Any subtype of it does not belong to the same result set because of $\prune$.
\end{itemize}
\saveSpaceFig

@@

The $\prune$ function takes a set of
types, and filters out those that have subtypes in the same set. In the returned set,
none of them has subtyping relation to one another, since all supertypes have been removed.
%%%%============================ prune(set) end ================%%%%%%%%

### 2017-10-20 NewVersion/sections/Discussion.tex
spellcheck, grammar

Abstract methods are one of the key features in most general OO languages. For example, Java interfaces were designed
to include only method declarations, and those abstract methods can be implemented in a class body. 
The formal Featherweight Java model~\cite{Igarashi01FJ} does not include abstract methods because of the orthogonality
to the core calculus. In traits the
similar idea is to use keywords like ``\textbf{require}'' for abstract method declarations~\cite{scharli03traits}.
Abstract methods provide a way to
delay the implementations to future subtypes. Using overriding, they also help to ``exclude'' existing implementations.

@@

Our formalization has a simple form of abstract methods, which behave similarly to conventional methods with respect to conflicts.
Other languages may support different behaviors.
For instance, in Java 8 when putting two identical abstract methods together by multiple inheritance, there is no conflict error. In Figure~\ref{fig:abstractdiamond}, we use italic $\emph{m}$ to denote abstract methods. In both cases, the Java compiler accepts the definition of $C$ and automatically merges the two inherited methods $m$ into a single one. \MIM{} behaves differently from Java in both cases.
In the triangle inheritance case (left) $C$ will have two distict abstract methods corresponding to $A.m$ and $B.m$. 
In the diamond inheritance case the definition of $C$ is rejected. 
There are two reasons for this difference in behaviour. Firstly, our formalization just treats abstract methods as concrete methods with an empty body, and that simplifies
the rules and proofs a lot. Secondly, and more importantly, we distinguish and treat differently conflicting methods, since they may represent different operations, even if they are abstract. Thus our model adopts a very conservative behavior rather than automatically merging 
methods by default (as done in many languages). 
Arguably, for the diamond case, as we have mentioned, it is actually an intentional conflict due to the same source $T$.
It is possible to change our model to account for other behaviors for abstract methods, but we view this as a mostly 
orthogonal change to our work, that should not affect the essence of the model presented here.

>>>>

Abstract methods are one of the key features in most general OO languages. For example, Java interfaces were designed
to include only method declarations, and those abstract methods can be implemented in a class body. 
The formal Featherweight Java model~\cite{Igarashi01FJ} does not include abstract methods because of the orthogonality
to the core calculus. In traits, the
similar idea is to use keywords like ``\textbf{require}'' for abstract method declarations~\cite{scharli03traits}.
Abstract methods provide a way to
delay the implementations to future subtypes. Using overriding, they also help to ``exclude'' existing implementations.

@@

Our formalization has a simple form of abstract methods, which behave similarly to conventional methods with respect to conflicts.
Other languages may support different behaviors.
For instance, in Java 8 when putting two identical abstract methods together by multiple inheritance, there is no conflict error. In Figure~\ref{fig:abstractdiamond}, we use italic $\emph{m}$ to denote abstract methods. In both cases, the Java compiler accepts the definition of $C$ and automatically merges the two inherited methods $m$ into a single one. \MIM{} behaves differently from Java in both cases.
In the triangle inheritance case (left) $C$ will have two distinct abstract methods corresponding to $A.m$ and $B.m$. 
In the diamond inheritance case, the definition of $C$ is rejected. 
There are two reasons for this difference in behaviour. Firstly, our formalization just treats abstract methods as concrete methods with an empty body, and that simplifies
the rules and proofs a lot. Secondly, and more importantly, we distinguish and treat differently conflicting methods, since they may represent different operations, even if they are abstract. Thus our model adopts a very conservative behavior rather than automatically merging 
methods by default (as done in many languages). 
Arguably, for the diamond case, as we have mentioned, it is actually an intentional conflict due to the same source $T$.
It is possible to change our model to account for other behaviors for abstract methods, but we view this as a mostly 
orthogonal change to our work, that should not affect the essence of the model presented here.

### 2017-10-20 NewVersion/sections/Formalization.tex
spellcheck, grammar

\textsc{(T-Invk)} is the typing rule for method invocation.
Naturally, the receiver and the arguments are required to be well-typed.
$\mbody$ is our key function for method lookup that implements the
hierarchical dispatching algorithm. The formal definition will be introduced in Section~\ref{sec:auxdefs}.
Here $\mbody(m, I_0, I_0)$ finds the most specific $m$ above $I_0$. ``Above $I_0$'' specifies
the search space, namely the super types of $I_0$ including itself.
For the general case, however, the hierarchical invocation $\mbody(m, I, J)$ finds ``the most specific $m$
above $I$ and along path/branch $J$''. ``Along path $J$'' additionally requires the result to relate to $J$, that is to say,
the most specific interface that has a subtyping relationship with $J$.

@@

\textsc{(T-Method)} is more interesting since a method can either be an original method or a hierarchical overriding, though
they share the same syntax and method typing rule. $\mostSpecific(m, I, J)$ is a fundamental function,
used to find ``the most specific interfaces that are above $I$ and
along path $J$, and originally defines $m$'' (see
Section~\ref{sec:auxdefs} for full definition).
By ``most specific interfaces'',
it implies that the inherited super types are excluded. Thus the condition $\mostSpecific(m, I, J) = \{J\}$ indicates a characteristic of a hierarchical overriding: it must override an original method; the overriding is direct and there does not exist any other original method $m$ in between.
Then $\mbody(m, J, J)$ provides the type of the original method, so hierarchical overriding has to preserve the type. Finally the return expression
is type-checked to be subtype of the declared return type. For the definition of an original method, $I$ equals $J$ and the rule is straightforward. \textsc{(T-AbsMethod)} is a similar rule but works on abstract method declarations.

\textsc{(T-Intf)} defines the typing rule on interfaces. The first condition is obvious, namely its methods need to be well checked. The third
condition checks whether the overriding between original methods preserves typing. In this condition we again use some helper functions defined in  Section~\ref{sec:auxdefs}. $I[m\ \kwoverride\ I]$ is defined if $I$ originally defines $m$, and $\canOverride(m, I, J)$ checks whether $I.m$ has the same type as $J.m$. Generally the preservation of method type is required for any super type $J$ and any method $m$.

@@

\paragraph{Semantic Rules} \textsc{(S-Invk)} is the only computation rule we need for method invocation.
As a small-step rule and by congruence, it assumes that the receiver and the arguments are already values.
Specifically, the receiver $(J)\new{I}$ indicates the dynamic type $I$
together with the static type $J$. Therefore $\mbody(m, I, J)$ carries out hierarchical dispatching, acquires
the types, the return expression $e_0$ and the interface $I_0$ which provides the most specific method.
Here we use $e_0$ to imply that the return expression is forced to be non-empty because it requires a concrete implementation. Now the
rule reduces method invocation to $e_0$ with substitution.
Parameters are substituted with arguments, and the \lstinline|this| reference is substituted by the receiver,
and in the meanwhile the static types are recorded via annotations. Finally the return type $I_e$ is put in the front as an annotation.
\paragraph{Congruence Rules} \textsc{(C-Receiver)}, \textsc{(C-Args)} and \textsc{(C-FReduce)} are natural congruence rules
on receivers, arguments, and cast-expressions, respectively. \textsc{(C-StaticType)} automatically adds an annotation $I$ to the new
object $\new{I}$. \textsc{(C-AnnoReduce)} merges nested upcasts into a single upcast with the outermost type.

>>>>

\textsc{(T-Invk)} is the typing rule for method invocation.
Naturally, the receiver and the arguments are required to be well-typed.
$\mbody$ is our key function for method lookup that implements the
hierarchical dispatching algorithm. The formal definition will be introduced in Section~\ref{sec:auxdefs}.
Here $\mbody(m, I_0, I_0)$ finds the most specific $m$ above $I_0$. ``Above $I_0$'' specifies
the search space, namely the supertypes of $I_0$ including itself.
For the general case, however, the hierarchical invocation $\mbody(m, I, J)$ finds ``the most specific $m$
above $I$ and along path/branch $J$''. ``Along path $J$'' additionally requires the result to relate to $J$, that is to say,
the most specific interface that has a subtyping relationship with $J$.

@@

\textsc{(T-Method)} is more interesting since a method can either be an original method or a hierarchical overriding, though
they share the same syntax and method typing rule. $\mostSpecific(m, I, J)$ is a fundamental function,
used to find ``the most specific interfaces that are above $I$ and
along path $J$, and originally defines $m$'' (see
Section~\ref{sec:auxdefs} for full definition).
By ``most specific interfaces'',
it implies that the inherited supertypes are excluded. Thus the condition $\mostSpecific(m, I, J) = \{J\}$ indicates a characteristic of a hierarchical overriding: it must override an original method; the overriding is direct and there does not exist any other original method $m$ in between.
Then $\mbody(m, J, J)$ provides the type of the original method, so hierarchical overriding has to preserve the type. Finally the return expression
is type-checked to be subtype of the declared return type. For the definition of an original method, $I$ equals $J$ and the rule is straightforward. \textsc{(T-AbsMethod)} is a similar rule but works on abstract method declarations.

\textsc{(T-Intf)} defines the typing rule on interfaces. The first condition is obvious, namely its methods need to be well checked. The third
condition checks whether the overriding between original methods preserves typing. In this condition we again use some helper functions defined in  Section~\ref{sec:auxdefs}. $I[m\ \kwoverride\ I]$ is defined if $I$ originally defines $m$, and $\canOverride(m, I, J)$ checks whether $I.m$ has the same type as $J.m$. Generally the preservation of method type is required for any supertype $J$ and any method $m$.

@@

\paragraph{Semantic Rules} \textsc{(S-Invk)} is the only computation rule we need for method invocation.
As a small-step rule and by congruence, it assumes that the receiver and the arguments are already values.
Specifically, the receiver $(J)\new{I}$ indicates the dynamic type $I$
together with the static type $J$. Therefore $\mbody(m, I, J)$ carries out hierarchical dispatching, acquires
the types, the return expression $e_0$ and the interface $I_0$ which provides the most specific method.
Here we use $e_0$ to imply that the return expression is forced to be non-empty because it requires a concrete implementation. Now the
rule reduces method invocation to $e_0$ with substitution.
Parameters are substituted with arguments, and the \lstinline|this| reference is substituted by the receiver,
and in the meanwhile the static types are recorded via annotations. Finally, the return type $I_e$ is put in the front as an annotation.
\paragraph{Congruence Rules} \textsc{(C-Receiver)}, \textsc{(C-Args)} and \textsc{(C-FReduce)} are natural congruence rules
on receivers, arguments, and cast-expressions, respectively. \textsc{(C-StaticType)} automatically adds an annotation $I$ to the new
object $\new{I}$. \textsc{(C-AnnoReduce)} merges nested upcasts into a single upcast with the outermost type.

### 2017-10-20 NewVersion/sections/Overview.tex
spellcheck, grammar

\begin{lstlisting}
interface Deck { void draw() {...} }
interface Drawable { void draw() {...} }
interface DrawableDeck extends Drawable, Deck {}
// main program
((Deck) new DrawableDeck()).draw()  // calls Deck.draw
// (new DrawableDeck()).draw()        // this call is ambiguous and rejected
\end{lstlisting}
In our language, the main program is merely an expression. The above cast indicates
that we expect to invoke the \lstinline|draw| method from the branch
\lstinline|Deck|. Similarly we could have used an upcast to \lstinline|Drawable|
to call the \lstinline|draw| method from \lstinline|Drawable|.
Without the cast, the call would be ambiguous and
\MIM{}'s type system would reject that call.

@@

\noindent Indeed, using the a reduction model similar to FJ~\cite{Igarashi01FJ} (without static types tracked), the reduction steps would roughly be:

@@

\paragraph{\MIM's solution:} 
The essential problem is how to ensure that the correct method is
invoked. To solve this problem, \MIM{} uses a variant of method
dispatching that we call \textit{hierarchical dispatching}. In
hierarchical dispatching both the static and dynamic type information 
are used to select the right method implementation.
%Inspired by C++, we have formalized this feature into a minimal core calculus,
%studied and proved the type soundness. It is concise and can readily be imported to other systems. 
%And in the formalization, we propose \textit{hierarchical dispatch} in our model
%as the default method lookup algorithm. 
During runtime, a method call
makes use of both the static type and the dynamic type of the
receiver, so it is a
combination of static and dynamic dispatching. Intuitively, the static type specifies one branch
to avoid ambiguity, and the dynamic type finds the most specific implementation on that branch.
To be specific, the following code is accepted by \MIM{}:
\begin{lstlisting}
interface Deck {
  void draw() {...}
  void shuffle() {...}
  void shuffleAndDraw() { this.shuffle(); this.draw(); }
}
interface Drawable {...}
interface SafeDeck extends Deck {...}
interface DrawableSafeDeck extends Drawable, SafeDeck {}
// main program
new DrawableSafeDeck().shuffleAndDraw() // SafeDeck.draw is called
\end{lstlisting}

@@

Notably, we track the static types by adding upcasts during reduction. In contrast to FJ,
where \lstinline|new C()| is a value, in \MIM{} such expression is not 
a value. Instead an expression of the form \lstinline|new C()| is reduced to
\lstinline|(C) new C()|, which is a value in \MIM{} and where the type
in the cast denotes the static type of the expression. This rule is
applied in the first reduction step. 
In the second reduction step, when 
\lstinline|shuffleAndDraw| is dispatched, the receiver
\lstinline|(DrawableSafeDeck) new DrawableSafeDeck()|
replaces the special variable \lstinline|this| by 
\lstinline|(Deck) new DrawableSafeDeck()|. Here, the static type used
in the cast (\lstinline|Deck|) denotes the origin of the \lstinline|shuffleAndDraw|
method, which is discovered during method lookup. 
%The static type information about the origin
%of the method is helpful for future disambiguation. 
Later, in the fourth step,
\lstinline|((Deck) new DrawableSafeDeck()).draw()|
is an instance of \textit{hierarchical invocation}, which can be read as 
``finding the most specific \lstinline|draw| above \lstinline|DrawableSafeDeck| and along path \lstinline|Deck|''. The meaning of ``above \lstinline|DrawableSafeDeck|'' implies its super types, and ``along path
\lstinline|Deck|'' specifies the branch. Finally, in the last
reduction step, we find the most
specific version of \lstinline|draw| in \lstinline|SafeDeck|. In this
sequence of
reduction steps, the cast that tracks the origin of
\lstinline|shuffleAndDraw| is crucial to unambiguously find the
correct implementation of \lstinline|draw|.
The formal procedure will be introduced in Section~\ref{sec:formalization} and Section~\ref{sec:auxdefs}.
%Note that in C++, if we put the \lstinline[language=c++]|virtual| keyword in front of every \lstinline|draw| method, that means
%we enforce the dynamic dispatch, and it does the disambiguation correctly.
%Unfortunately, the C++ language, or rather the feature on triangle inheritance, has not been well formalized.
%In our language model, there are two types of casts: \emph{implicit casts} and \emph{explicit upcasts}. The implicit casts appear
%when passing arguments and returning expressions, they are also
%upcasts. Hence all casts are safe. \bruno{This discussion seems
%  inconplete. Either complete it or remove it.}

@@

With unintentional method conflicts, however, the situation is more
complicated because different, separate, conflicting methods can coexist
in one class. Ideally we would like to support overriding for those
methods too, in exactly the same way that overriding is available for 
other (non-conflicting) methods. However we need to be able to
override the individual conflicting methods, rather than overriding all
conflicting methods into a single merged one. We illustrate the
problem and the need for a more refined overriding mechanism with 
an example.

Suppose that the programmer defines a new interface \lstinline|DrawableSafeDeck| (based on the code in Section~\ref{subsec:problem2} without the old \lstinline|DrawableSafeDeck|), but he needs to override
\lstinline|Drawable.draw| and give a new implementation of drawing, so that the deck can indeed be visualized on the canvas.

@@

\paragraph{Potential solutions/workarounds in existing languages:} Unfortunately in all languages we have known
(including C++), the existing approaches are unsatisfactory.
One direction is to simply avoid this issue, by putting overriding before inheritance. For example, as shown in Figure~\ref{fig:drawablesafedeck}(middle), we define a new component \lstinline|DrawableRect| that extends \lstinline|Drawable|, which simply draws the deck as a rectangle, and modifies the hierarchy:
\begin{lstlisting}
interface DrawableRect extends Drawable {
  void draw() {
    JFrame frame = new JFrame("Canvas");
    frame.setSize(600, 600);
    frame.getContentPane().setBackground(Color.red);
    frame.getContentPane().add(new Square(10,10,100,100)); ...
  }
}
interface DrawableSafeDeck extends DrawableRect, SafeDeck {}
\end{lstlisting}
This workaround seems to work, but there are severe issues:
\begin{itemize}
	\item It changes the hierarchy and existing code, hence breaks the modularity.
	\item Separate overriding is required to come after the triangle
          inheritance, especially when the implementation needs functionality from both parents. In the above code we have assumed
	that the overriding is unrelated to \lstinline|Deck|. But when the drawing relies on some information of the \lstinline|Deck|
	object, we have to either introduce field methods for delegation, or change the signature of \lstinline|draw| to take a parameter.
	Either way introduces unnecessary complexity and affects extensibility.
\end{itemize}

@@

\paragraph{\MIM{}'s solution:} An additional feature of our model is \textit{hierarchical overriding}. It allows conflicting methods
to be overridden on individual branches, and hence offers independent extensibility. The above example can be easily realized by:
\begin{lstlisting}
interface DrawableSafeDeck extends Drawable, SafeDeck {
  void draw() override Drawable {
    JFrame frame = new JFrame("Canvas");
    frame.setSize(600, 600);
    frame.getContentPane().setBackground(Color.red);
    frame.getContentPane().add(new Square(10,10,100,100)); ...
  }
}
// main program
((Drawable)new DrawableSafeDeck()).draw(); // calls the latest draw
\end{lstlisting}

@@

\paragraph{Terminology:} In \lstinline|Drawable|, \lstinline|Deck| and \lstinline|SafeDeck|, the \lstinline|draw| methods are called \textit{original methods} in this paper, because they are originally defined by the interfaces.
In contrast, \lstinline|DrawableSafeDeck| defines a \textit{hierarchical overriding method}. The difference is that traditional method overriding overrides all branches by defining another original method, whereas hierarchical overriding only refines one branch.

@@

In  Figure~\ref{fig:examplesmbody}, (a) is the base case for unintentional conflicts, namely the triangle inheritance. (b) uses overriding to explicitly merge the conflicting methods. (c) represents hierarchical overriding. (d) and (e) are two base cases of diamond inheritance in \MIM{}, and the definition of each $C$ is rejected because $T$ is an ambiguous parent to $C$. (f) gives a common solution to the diamond as in Java or traits, which is to explicitly override $A.m$ and $B.m$ in $C$. In the last three examples, conflicting methods $A.m$ and $B.m$ should be viewed as
intentional conflicts, as they come from the same source $T$.
% Similar to many OO languages, the model also allows \textit{super method invocation} in a method body. The invocation \lstinline|super.T::m()| will ignore all the subtypes of \lstinline|T|, and only look at \lstinline|T| together with its super interfaces. It should behave the same as \lstinline|new T().m()| in principle.

>>>>

\begin{lstlisting}
interface Deck { void draw() {...} }
interface Drawable { void draw() {...} }
interface DrawableDeck extends Drawable, Deck {}
// main program
((Deck) new DrawableDeck()).draw()  // calls Deck.draw
// (new DrawableDeck()).draw()        // this call is ambiguous and rejected
\end{lstlisting}
In our language, the main program is merely an expression. The above cast indicates
that we expect to invoke the \lstinline|draw| method from the branch
\lstinline|Deck|. Similarly, we could have used an upcast to \lstinline|Drawable|
to call the \lstinline|draw| method from \lstinline|Drawable|.
Without the cast, the call would be ambiguous and
\MIM{}'s type system would reject it.

@@

\noindent Indeed, using a reduction model similar to FJ~\cite{Igarashi01FJ} (without static types tracked), the reduction steps would roughly be:

@@

\paragraph{\MIM's solution:} 
The essential problem is how to ensure that the correct method is
invoked. To solve this problem, \MIM{} uses a variant of method
dispatching that we call \textit{hierarchical dispatching}. In
hierarchical dispatching, both the static and dynamic type information 
are used to select the right method implementation.
%Inspired by C++, we have formalized this feature into a minimal core calculus,
%studied and proved the type soundness. It is concise and can readily be imported to other systems. 
%And in the formalization, we propose \textit{hierarchical dispatch} in our model
%as the default method lookup algorithm. 
During runtime, a method call
makes use of both the static type and the dynamic type of the
receiver, so it is a
combination of static and dynamic dispatching. Intuitively, the static type specifies one branch
to avoid ambiguity, and the dynamic type finds the most specific implementation on that branch.
To be specific, the following code is accepted by \MIM{}:
\begin{lstlisting}
interface Deck {
  void draw() {...}
  void shuffle() {...}
  void shuffleAndDraw() { this.shuffle(); this.draw(); }
}
interface Drawable {...}
interface SafeDeck extends Deck {...}
interface DrawableSafeDeck extends Drawable, SafeDeck {}
// main program
new DrawableSafeDeck().shuffleAndDraw() // SafeDeck.draw is called
\end{lstlisting}

@@

Notably, we track the static types by adding upcasts during reduction. In contrast to FJ,
where \lstinline|new C()| is a value, in \MIM{} such expression is not 
a value. Instead, an expression of the form \lstinline|new C()| is reduced to
\lstinline|(C) new C()|, which is a value in \MIM{} and where the type
in the cast denotes the static type of the expression. This rule is
applied in the first reduction step. 
In the second reduction step, when 
\lstinline|shuffleAndDraw| is dispatched, the receiver
\lstinline|(DrawableSafeDeck) new DrawableSafeDeck()|
replaces the special variable \lstinline|this| by 
\lstinline|(Deck) new DrawableSafeDeck()|. Here, the static type used
in the cast (\lstinline|Deck|) denotes the origin of the \lstinline|shuffleAndDraw|
method, which is discovered during method lookup. 
%The static type information about the origin
%of the method is helpful for future disambiguation. 
Later, in the fourth step,
\lstinline|((Deck) new DrawableSafeDeck()).draw()|
is an instance of \textit{hierarchical invocation}, which can be read as 
``finding the most specific \lstinline|draw| above \lstinline|DrawableSafeDeck| and along path \lstinline|Deck|''. The meaning of ``above \lstinline|DrawableSafeDeck|'' implies its supertypes, and ``along path
\lstinline|Deck|'' specifies the branch. Finally, in the last
reduction step, we find the most
specific version of \lstinline|draw| in \lstinline|SafeDeck|. In this
sequence of
reduction steps, the cast that tracks the origin of
\lstinline|shuffleAndDraw| is crucial to unambiguously find the
correct implementation of \lstinline|draw|.
The formal procedure will be introduced in Section~\ref{sec:formalization} and Section~\ref{sec:auxdefs}.
%Note that in C++, if we put the \lstinline[language=c++]|virtual| keyword in front of every \lstinline|draw| method, that means
%we enforce the dynamic dispatch, and it does the disambiguation correctly.
%Unfortunately, the C++ language, or rather the feature on triangle inheritance, has not been well formalized.
%In our language model, there are two types of casts: \emph{implicit casts} and \emph{explicit upcasts}. The implicit casts appear
%when passing arguments and returning expressions, they are also
%upcasts. Hence all casts are safe. \bruno{This discussion seems
%  inconplete. Either complete it or remove it.}

@@

With unintentional method conflicts, however, the situation is more
complicated because different, separate, conflicting methods can coexist
in one class. Ideally, we would like to support overriding for those
methods too, in exactly the same way that overriding is available for 
other (non-conflicting) methods. However, we need to be able to
override the individual conflicting methods, rather than overriding all
conflicting methods into a single merged one. We illustrate the
problem and the need for a more refined overriding mechanism with 
an example.

Suppose that the programmer defines a new interface \lstinline|DrawableSafeDeck| (based on the code in Section~\ref{subsec:problem2} without the old \lstinline|DrawableSafeDeck|), but he needs to override
\lstinline|Drawable.draw| and give a new implementation of drawing so that the deck can indeed be visualized on the canvas.

@@

\paragraph{Potential solutions/workarounds in existing languages:} Unfortunately in all languages we have known
(including C++), the existing approaches are unsatisfactory.
One direction is to simply avoid this issue, by putting overriding before inheritance. For example, as shown in Figure~\ref{fig:drawablesafedeck}(middle), we define a new component \lstinline|DrawableRect| that extends \lstinline|Drawable|, which simply draws the deck as a rectangle, and modifies the hierarchy:
\begin{lstlisting}
interface DrawableRect extends Drawable {
  void draw() {
    JFrame frame = new JFrame("Canvas");
    frame.setSize(600, 600);
    frame.getContentPane().setBackground(Color.red);
    frame.getContentPane().add(new Square(10,10,100,100)); ...
  }
}
interface DrawableSafeDeck extends DrawableRect, SafeDeck {}
\end{lstlisting}
This workaround seems to work, but there are severe issues:
\begin{itemize}
	\item It changes the hierarchy and existing code, hence breaks the modularity.
	\item Separate overriding is required to come after the triangle
          inheritance, especially when the implementation needs functionality from both parents. In the above code, we have assumed
	that the overriding is unrelated to \lstinline|Deck|. But when the drawing relies on some information of the \lstinline|Deck|
	object, we have to either introduce field methods for delegation or change the signature of \lstinline|draw| to take a parameter.
	Either way introduces unnecessary complexity and affects extensibility.
\end{itemize}

@@

\paragraph{\MIM{}'s solution:} An additional feature of our model is \textit{hierarchical overriding}. It allows conflicting methods
to be overridden on individual branches, hence offers independent extensibility. The above example can be easily realized by:
\begin{lstlisting}
interface DrawableSafeDeck extends Drawable, SafeDeck {
  void draw() override Drawable {
    JFrame frame = new JFrame("Canvas");
    frame.setSize(600, 600);
    frame.getContentPane().setBackground(Color.red);
    frame.getContentPane().add(new Square(10,10,100,100)); ...
  }
}
// main program
((Drawable)new DrawableSafeDeck()).draw(); // calls the latest draw
\end{lstlisting}

@@

\paragraph{Terminology:} In \lstinline|Drawable|, \lstinline|Deck|, and \lstinline|SafeDeck|, the \lstinline|draw| methods are called \textit{original methods} in this paper, because they are originally defined by the interfaces.
In contrast, \lstinline|DrawableSafeDeck| defines a \textit{hierarchical overriding method}. The difference is that traditional method overriding overrides all branches by defining another original method, whereas hierarchical overriding only refines one branch.

@@

In  Figure~\ref{fig:examplesmbody}, (a) is the base case for unintentional conflicts, namely the triangle inheritance. (b) uses overriding to explicitly merge the conflicting methods. (c) represents hierarchical overriding. (d) and (e) are two base cases of diamond inheritance in \MIM{} and the definition of each $C$ is rejected because $T$ is an ambiguous parent to $C$. (f) gives a common solution to the diamond as in Java or traits, which is to explicitly override $A.m$ and $B.m$ in $C$. In the last three examples, conflicting methods $A.m$ and $B.m$ should be viewed as
intentional conflicts, as they come from the same source $T$.
% Similar to many OO languages, the model also allows \textit{super method invocation} in a method body. The invocation \lstinline|super.T::m()| will ignore all the subtypes of \lstinline|T|, and only look at \lstinline|T| together with its super interfaces. It should behave the same as \lstinline|new T().m()| in principle.

### 2017-10-20 NewVersion/sections/RelatedWork.tex
spellcheck, grammar

Class \texttt{MiddleMan}, as its name suggests, acts as a middle man between its class \texttt{C} and its parents \texttt{A, B}. \texttt{MiddleMan} defines a virtual method \texttt{m} that overrides a parent method \texttt{m} and delegates the implementation to another method \texttt{m\_impl} that takes $\kwthis$ as parameter. C++ supports method overloading, so that multiple \texttt{m\_impl} methods with different parameter types can coexist. When defining class \texttt{C}, we specify the parents to be \texttt{MiddleMan<A>, MiddleMan<B>} instead of \texttt{A, B}. In this way, programmers may define new versions of \texttt{A.m} and \texttt{B.m} in class \texttt{C} by providing the corresponding \texttt{m\_impl} methods. Then in the client code, the method call \texttt{((A*)c)->m()} will print out string \texttt{"MA2"}, as expected. Although this workaround can help us defining partial method overrides to a certain extent, the drawbacks are obvious. Firstly, the approach is complex and requires the pre-knowledge of the programmer to this approach. Moreover, the lack of direct syntax support makes MiddleMan code cumbersome to write. Finally, the approach is ad-hoc, meaning that the class \texttt{MiddleMan} shown in Figure~\ref{fig:middleman} is not general enough to be used in other cases: more middle-mans are needed if partial method overrides happens in other classes; and it is even worse when return types differ.

>>>>

Class \texttt{MiddleMan}, as its name suggests, acts as a middle man between its class \texttt{C} and its parents \texttt{A, B}. \texttt{MiddleMan} defines a virtual method \texttt{m} that overrides a parent method \texttt{m} and delegates the implementation to another method \texttt{m\_impl} that takes $\kwthis$ as a parameter. C++ supports method overloading, so that multiple \texttt{m\_impl} methods with different parameter types can coexist. When defining class \texttt{C}, we specify the parents to be \texttt{MiddleMan<A>, MiddleMan<B>} instead of \texttt{A, B}. In this way, programmers may define new versions of \texttt{A.m} and \texttt{B.m} in class \texttt{C} by providing the corresponding \texttt{m\_impl} methods. Then in the client code, the method call \texttt{((A*)c)->m()} will print out string \texttt{"MA2"}, as expected. Although this workaround can help us defining partial method overrides to a certain extent, the drawbacks are obvious. Firstly, the approach is complex and requires the pre-knowledge of the programmer to this approach. Moreover, the lack of direct syntax support makes MiddleMan code cumbersome to write. Finally, the approach is ad-hoc, meaning that the class \texttt{MiddleMan} shown in Figure~\ref{fig:middleman} is not general enough to be used in other cases: more middle-mans are needed if partial method overrides happens in other classes; and it is even worse when return types differ.
